---
title: "PD_exploration"
author: "Pascal"
date: "2022-12-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 2

### a) paramters for KD adb receotir occupancy RO

```{r}
setwd("~/BSc_UZH/UZH_22HS/BIO445_BK/12_PK_PD_Stochiometry")

E <- function(conc, E0, Emax, EC50, gamma) {
  
  tot_eff <- E0 + ((Emax * conc ** gamma) / (EC50 ** gamma + conc ** gamma))
    
  return(tot_eff)
}
```

### b) minimum efficacy and potency

```{r}
Emax <- 90 # maximum treatment effect
E0 <- 10 # baseline effect
EC50 <- 0.3 # 50 % of drug effect
# Hill coefficient
gammaA <- 1
gammaB <- 4
gammaC <- 0.25
conc <- seq(0, 1.5, by = 0.015)

drug_A <- E(conc, E0, Emax, EC50, gammaA) # possible to directly plug in without for loop
drug_B <- E(conc, E0, Emax, EC50, gammaB)
drug_C <- E(conc, E0, Emax, EC50, gammaC)

drug_frame <- data.frame(values = c(drug_A, drug_B, drug_C),
                         conc = rep(conc, 3),
                         drug = c(rep("drug_A", length(drug_A)), 
                                  rep("drug_B", length(drug_B)),
                                  rep("drug_C", length(drug_C))))
```

#### ggplot

```{r}
library(ggplot2)
library(ggforce)
library(tidyverse)
           
p <- ggplot(data = drug_frame, mapping = aes(x = conc, y = values, color = drug)) + 
  geom_line() + 
  geom_vline(xintercept = EC50, linetype = "dashed") + 
  labs(x = "concentration", y = "effect size", title = "Sigmoid curves of different drugs") + 
  theme_bw()

# required: minimum therapeutic effect, c > too toxic

min_ther <- 20 # min therapeutic effect
max_c <- 0.25 # max concentration that can be taken

p + ggforce::facet_zoom(xlim = c(0, max_c), ylim = c(min_ther, 100)) +
  labs(title = "Drug C is best for (i)")

ggsave("2_b_i_drug_C.png", height = 10, width = 20, units = "cm")

ther_window <- c(0.45, 1.2) # wished therapeutic conc

p + ggforce::facet_zoom(xlim = ther_window, ylim = c(min_ther, 100)) + 
  labs(title = "Drug B is best for (ii)")

ggsave("2_b_i_drug_B.png", height = 10, width = 20, units = "cm")

```

# Problem 2: direct response and biophase distribution models

### a) PK model

```{r}
source("PD_scripts.r")

some_frame <- as.data.frame(conc_Blood_Brain())
exploratory_frame <- pivot_longer(data = some_frame,
                                  cols = c(Blood, Brain),
                                  names_to = "Type",
                                  values_to = "C")

ggplot(data = exploratory_frame, 
       mapping = aes(x = Time, y = C, color = Type)) + 
  geom_line() + 
  geom_vline(xintercept = 75, linetype = "dashed")
  labs(x = "Time [min]", y = "Concentration", 
       title = "Drug concetration and blood brain barrier") + 
  theme_bw()
  
ggsave("3_Exploratory_concBloodBrain_standard_params.png", 
       height = 10, width = 20, units = "cm")

# for quick exploration
plot(conc_Blood_Brain())
```

### b) direct response model

```{r}

E0 <- 0
Emax <- 100
EC50 <- 1
gamma <- 2


out <- conc_Blood_Brain()

effect_brain <- E(out_frame[,"Brain"],E0, Emax, EC50, gamma)

# determines highest effect
max_EGG <- which.max(effect_brain)

ggplot(mapping = aes(x = out[,"Time"], y = effect_brain)) + 
  geom_vline(xintercept = conc_brain[max_EGG], color = "red") +
  geom_line() + 
  labs(x = "Time [min]", y = "EEG", title = "Effect vs Time") +
  theme_bw()

# mask_1.5 <- out[,"Brain"] < 1.5

ggplot(mapping = aes(x = out[, "Brain"], y = effect_brain)) + 
  geom_vline(xintercept = conc_brain[max_EGG], color = "red") +
  geom_line() +
  xlim(0, 1.5) +
  labs(x = "c(Drug)", y = "EEG", title = "Effect vs Concentration") +
  theme_bw()

# max in EEG in red


```


### c) biophase distrubution model

Predict effects on brain by measuring effects in blood, because it is much less invasive.

```{r}

E0 <- 0
Emax <- 100
EC50 <- 1
gamma <- 2


out <- conc_Blood_Brain()

effect_brain <- E(out_frame[,"Blood"],E0, Emax, EC50, gamma)

ggplot(mapping = aes(x = out[,"Time"], y = effect_brain)) + 
  geom_vline(xintercept = conc_brain[max_EGG], color = "red") +
  geom_line() + 
  labs(x = "Time [min]", y = "EEG", title = "Effect vs Time") +
  theme_bw()

ggplot(mapping = aes(x = out[,"Blood"], y = effect_brain)) + 
  geom_vline(xintercept = conc_brain[max_EGG], color = "red") +
  geom_line() + 
  labs(x = "Concentration", y = "EEG", title = "Effect Brain vs Concentration Blood") +
  xlim(0, 3) +
  theme_bw()



```

### our slide: comparing the direct and biophase distribution model

```{r}
E0 <- 0
Emax <- 100
EC50 <- 0.3
gamma <- 2


out <- conc_Blood_Brain()

effect_brain <- E(out[,"Brain"],E0, Emax, EC50, gamma)

# determines highest effect
max_EGG <- which.max(effect_brain)

len <- length(out[, "Blood"])

df <- data.frame(effect = rep(effect_brain,2), 
                 conc = c(out[, "Brain"], out[,"Blood"]),
                 Model = c(rep("Direct", len), rep("Biophase", len)), 
                 Time = rep(out[,"Time"], 2))


p1 <- ggplot(data = df, mapping = aes(x = conc, y = effect)) +
  geom_line(aes(linetype = Model), linewidth = 1, alpha = 0.5) +
  geom_point(aes(color = Time)) +
  labs(x = "Drug concentration", 
       y = "EGG (effect)", 
       title = "Effect in brain vs. drug concentration", 
       subtitle = "Biophase distribution model displays accurate approximation of effect in brain") +
  theme_bw()

some_frame <- as.data.frame(out)
exploratory_frame <- pivot_longer(data = some_frame,
                                  cols = c(Blood, Brain),
                                  names_to = "Type",
                                  values_to = "C")

p2 <- ggplot(data = exploratory_frame, 
       mapping = aes(x = Time, y = C, color = Type)) + 
  geom_line() + 
  geom_vline(xintercept = c(17, 75), linetype = "dashed", alpha = 0.5) +
  labs(x = "Time [min]", y = "Drug concentration", 
       title = "Drug overcoming blood brain barrier", 
       subtitle = "Both tissues show similar concentration trajectories after equilibrium is reached.") + 
  theme_bw()


library(ggpubr)
ggarrange(p2, p1, ncol = 2, nrow = 1)

ggsave("Blood_brain_barrier_model.png", height = 10, width = 30, units = "cm")

```